---
output: html_document
---
#SATATISTICAL INFERENCE 

=====================================================================

**COURSE PROJECT**

**AUTHOR**: *Niane A. Demba*

------------------------

## Introduction
The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also also 1/lambda. In this simulation, I will investigate the distribution of averages of 40 exponential(0.2)s. 

### SIMULATIONS (part 1)
1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.

```{r loding data, echo=FALSE}
#Simulation of exponential distribution
n<-40; nosim <- 1000; lambda <- 0.2
#Data 
dist <- matrix(rexp(nosim*n, lambda), nosim)
data <- data.frame(dist)
```

 

```{r compute mean , echo=FALSE}
#Generate Distribution of mean columns data
meanDist <- vector(); sdDist <- vector()
for (i in 1:nrow(data)){
    meanDist[i] <- mean(dist[i,],na.rm=T)
    sdDist[i] <- sd(dist[i,],na.rm=T)
}  
new_data <- data.frame(cbind(meanDist,sdDist))
names(new_data) <- c("Mean","SD")
# compare means 
Table = data.frame(cbind(mean(new_data$Mean),1/lambda)); names(Table) <- c("Simulated","Theorical")
```

 

```{r compare , echo=FALSE, results = "asis"}
paste("The simulated center of the distribution ", Table$Simulated, "is close to  the theorical center of the distribution :", Table$Theorical)
```
 

```{r plots, echo=FALSE}
par(mfrow = c(1, 2), mar = c(4, 4, 2, 1), oma = c(0, 0, 2, 0)) 
with(data, {
    hist(new_data$Mean,main="Center of distribution")
    abline(v=mean(new_data$Mean),lwd=4,col="blue")
    hist(new_data$Mean,main="Theorical center")
    abline(v=1/lambda,lwd=4,col="green")
    mtext("Compare simulated center and theorical center of the exponential distribution", outer = TRUE)
})
```


2. Show how variable it is and compare it to the theoretical variance of the distribution.


```{r create table, echo=FALSE}
# compare means 
Table = data.frame(cbind(mean(new_data$SD),1/lambda));names(Table) <- c("Simulated","Theorical")
```

The table below shows compared variances
```{r show table, echo=FALSE, results = "asis"}
library(xtable);xt <- xtable(Table);print(xt, type = "html")
```

3. Show that the distribution is approximately normal.
The previous histogram of the distribution shows that the distribution is approximately normal

4. The coverage of the confidence interval for 1/lambda: X¯±1.96Sn√. (This only needs to be done for the specific value of lambda). 

```{r conf interval , echo=FALSE}
#Confidence interval  
xbar <- mean(new_data$Mean)
var <- var(new_data$Mean)
ME <- 1.96*sqrt(var/1000)
xbar +c(-ME,+ME)
```

### SIMULATIONS (part 2)
1. Load the ToothGrowth data and perform some basic exploratory data analyses 


```{r ToothGrowth data, echo=FALSE}
#Simulation of exponential distribution
library(datasets);data(ToothGrowth) 
data <- ToothGrowth
boxplot(data,main="ToothGrowth Box plot")
```

2. Basic summary of the data. 


```{r summmary data, echo=FALSE}
#Simulation of exponential distribution
summary(data)
```

3. Confidence intervals and hypothesis tests to compare tooth growth by supp and dose. 
```{r suppdose, echo=FALSE}
#Simulation of exponential distribution
#library(UsingR)
#Len for VC
vc05 = data$len[data$supp=="VC" & data$dose==0.5]; vc1 = data$len[data$supp=="VC" & data$dose==1.0];vc2 = data$len[data$supp=="VC" & data$dose==2.0]
#Len for OJ
oj05 = data$len[data$supp=="OJ" & data$dose==0.5]; oj1 = data$len[data$supp=="OJ" & data$dose==1.0];oj2 = data$len[data$supp=="OJ" & data$dose==2.0]
```

Confidence intervals and hypothesis tests to compare tooth growth by supp = VC and dose = 0.5 with supp = VC and dose = 1.0")

```{r suppdose test, echo=FALSE}
t.test(vc05 - vc1)

```

4. State your conclusions and the assumptions needed for your conclusions. 

**Conclusion**: Confidence intervals and hypothesis tests to compare tooth growth by supp = VC and dose = 0.5 with supp = VC and dose = 1.0 shows a small p-value so means are differents. 